---
title: "Propensity Score Weighting"
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
---

Propensity score weighting is one of the techniques used in controlling
for selection biases in nonexperimental studies. Propensity scores can
be used as weights to account for selection assignment differences
between treatment and comparison groups. One of the advantages of this
approach is that all the individuals in the study can be used for the
outcomes evaluation.

In order to conduct an analysis involving propensity scores, the authors
follow a very specific set of steps that include:

1.  Outcome analysis without the use of propensity scores

2.  Balance analysis prior to the implementation of propensity scores

3.  Propensity score estimation

4.  Weight estimation using propensity scores

5.  Balance analysis after implementing propensity scores

6.  Outcomes analysis using propensity scores in a weighted regression

We will use the "lalonde" datafile that is part of the {MatchIt} pckage. 
The dataset was developed by Lalonde (1986) to demonstrate the impact of a retraining program (National Supported Work Demonstration). The data file included in {MatchIt} contains 614 observations, with 185 in the treatment group and 429 in the control group. The outcome variable is re78, which is the income for individuals in both groups during 1978.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

```{r, include=FALSE}
#### Load the packages that are required.

###  Create a vector of packages to be installed
pkgs <- c("tidyverse","MatchIt", "broom", "weights")

## tidyverse: Data manipulation
## MatchIt: Match things

###  Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

###  Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

###  Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))
library(DDSQLtools)

###  Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

set.seed(1234)
```

#### 0. Read in the data

```{r}
attach(lalonde)
lalonde <- lalonde %>% 
  mutate(black = ifelse(race == "black", 1, 0),
         hispan = ifelse(race == "hispan", 1, 0))
```

#### 1.  Outcome analysis without the use of propensity scores

In this step, we run an outcome analysis without the use of propensity scores. 

This analysis is helpful to gauge what might have been the result of the outcome analysis had we not used propensity scores to control for potential selection biases associated with group assignment. 

```{r}
model0 <- lm(re78 ~ treat + black + hispan + married, data = lalonde)
summary(model0)
```

From the results shown above, there is no treatment effect (since the variable treat is not statistically significant), but the `black` and `married` covariates are significant at p < 0.05

#### 2.  Balance analysis prior to the implementation of propensity scores

This step is intended to assess the degree of bias between the groups before the propensity score is incorporated in the analysis. 

Typical analyses include statistical comparisons between the covariate (as DV) and the treatment variable (as IV).

```{r}
yvars <- names(lalonde)[!names(lalonde) %in% c("treat", "race")]
xvar <- "treat"

for(i in 1: length(yvars)){
form <- as.formula(paste(yvars[i], xvar, sep = "~"))
print(form)
print(t.test(form, data = lalonde))
  
}
```

The results show us that there is imbalance in treatment assignment in all the variables expect `married`.

#### 3.  Propensity score estimation

The next step is the estimation of the propensity scores that will be used as weights in the analysis. 

Logistic regression is used to determine the probability of membership in the treatment or control group, given the specific set of selection variables included. 
```{r}
## Fit the logistic regression model
ps <- glm(treat ~ age + educ + nodegree + re74 + re75, 
          family = binomial(), data = lalonde)

## Extract the predicted values (propensity scores) and attach them to the data
lalonde2 <- augment_columns(ps,lalonde,type.predict = "response") %>%
            rename(propensity = .fitted)
```

#### 4.  Weight estimation using propensity scores

Calculate the inverse probability weights by using the formula:

$$\frac{Treatment}{Propensity} + \frac{1-Treatment}{1-Propensity}$$

This gives more weight to the misclassified observations and less weight to the classifications that are classified correctly.

```{r}
lalonde2 <- lalonde2 %>% 
  mutate(ipw = (treat / propensity) + ((1 - treat)/(1 - propensity)))
```

#### 5.  Balance analysis after implementing propensity scores
The ultimate purpose of using propensity scores is to balance the treatment/comparison groups on the observed covariates.

To assess the success of the propensity scores as weights in a weighted regression for removing selection bias, a new set of tests to check the balance should be performed.

However, now weighted linear regressions are performed using computed propensity scores as weights. 

```{r}
yvars <- names(lalonde)[!names(lalonde) %in% c("treat", "race")]
xvar <- "treat"

for(i in 1: length(yvars)){
form <- as.formula(paste(yvars[i], xvar, sep = "~"))
print(form)
print(summary(lm(form, weight = (ipw),  data = lalonde2)))
  
}
```

#### 6.  Outcomes analysis using propensity scores in a weighted regression

The final step in the analysis is to run the outcomes model using the propensity scores as weights. In this example the outcomes model includes
re78 as the outcome variable and treat, black and married as independent variables. 

```{r}
model3 <- lm(re78 ~ treat + black + hispan + married, data = lalonde2,
             weights = (ipw))
summary(model3)
```
The results above show that the effect of the treatment was statistically significant after the groups are balanced using propensity scores. 
