---
title: "Randomization and Matching"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

```{r, include=FALSE}
#### Load the packages that are required.

###  Create a vector of packages to be installed
pkgs <- c("tidyverse", "ggdag", "scales", "broom", "patchwork")

## tidyverse: Data manipulation
## ggdag: Making DAGS 
## scales: Format numbers with functions like comma(), percent(), and dollar()
## broom: Convert models to data frames
## patchwork: Combine ggplots into single composite plots

###  Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

###  Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

###  Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))
library(DDSQLtools)

###  Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

set.seed(1234)
```


### Program details

In this hypothetical situation, an NGO is planning on launching a training program designed to boost incomes. Based on their experiences in running pilot programs in other countries, they’ve found that older, richer men tend to self-select into the training program. The NGO’s evaluation consultant (you!) drew this causal model explaining the effect of the program on participant incomes, given the confounding caused by age, sex, and prior income:

```{r}
village_randomized <- read_csv("data/village_randomized.csv")
```


```{r}
income_dag <- dagify(post_income ~ program + age + sex + pre_income,
                     program ~ age + sex + pre_income,
                     exposure = "program",
                     outcome = "post_income",
                     labels = c(post_income = "Post income",
                                program = "Program",
                                age = "Age",
                                sex = "Sex",
                                pre_income = "Pre income"),
                     coords = list(x = c(program = 1, post_income = 5, age = 2,
                                         sex = 4, pre_income = 3),
                                   y = c(program = 2, post_income = 2, age = 1,
                                         sex = 1, pre_income = 3)))

ggdag_status(income_dag, use_labels = "label", text = FALSE, seed = 1234) +
  guides(color = FALSE) +
  theme_dag()
```

The NGO just received funding to run a randomized controlled trial (RCT) in a village, and you’re excited because you can finally manipulate access to the program—you can calculate $E(Postincome|do(Program))$ . Following the rules of causal diagrams, you get to delete all the arrows going into the program node:

```{r}
income_dag_rct <- dagify(post_income ~ program + age + sex + pre_income,
                         exposure = "program",
                         outcome = "post_income",
                         labels = c(post_income = "Post income",
                                    program = "Program",
                                    age = "Age",
                                    sex = "Sex",
                                    pre_income = "Pre income"),
                         coords = list(x = c(program = 1, post_income = 5, age = 2,
                                             sex = 4, pre_income = 3),
                                       y = c(program = 2, post_income = 2, age = 1,
                                             sex = 1, pre_income = 3)))

ggdag_status(income_dag_rct, use_labels = "label", text = FALSE, seed = 1234) +
  guides(color = FALSE) +
  theme_dag()
```

There are two steps involved while analysing RCTS,

+ Check that key demographics and other confounders are balanced.

+ Find differences in average outcome in treatment and control groups


### 1. Check that key demographics and other confounders are balanced.
You ran the study on 1,000 participants over the course of 6 months and you just got your data back. Before calculating the effect of the program, you first check to see how well balanced assignment was

```{r}
village_randomized %>%
  count(program) %>%
  mutate(prop = n / sum(n))
```
You then check to see how well balanced the treatment and control groups were in participants' pre-treatment characteristics:

```{r}
village_randomized %>%
  group_by(program) %>%
  summarize(prop_male = mean(sex_num),
            avg_age = mean(age),
            avg_pre_income = mean(pre_income))
```

These variables appear fairly well balanced. To check that there aren’t any statistically significant differences between the groups, you make some graphs and/or run t-tests

#### 1.1 sex

```{r}
## t-test
## Null hypothesis: no difference
t.test(sex_num ~ program, data = village_randomized)

## graph
plot_diff_sex <- ggplot(village_randomized, aes(x = program, y = sex_num, color = program)) +
  stat_summary(geom = "pointrange", fun.data = "mean_se", fun.args = list(mult = 1.96)) +
  guides(color = FALSE) +
  scale_color_manual(values = c("maroon", "purple"))+
  labs(x = NULL, y = "Proportion male")
plot_diff_sex
```

There is no significant difference in sex distribution between those who received the training program (treatment group) and those who did not (control group).  

#### 1.2 age

```{r}
## t-test
## Null hypothesis: no difference
t.test(age ~ program, data = village_randomized)

## graph
plot_diff_age <- ggplot(village_randomized, aes(x = program, y = age, color = program)) +
  stat_summary(geom = "pointrange", fun.data = "mean_se", fun.args = list(mult = 1.96)) +
  guides(color = FALSE) +
  scale_color_manual(values = c("maroon", "purple"))+
  labs(x = NULL, y = "age")

plot_diff_age
```

There is no significant difference in age between those who received the training program (treatment group) and those who did not (control group).  

#### 1.3 pre-income

```{r}
## t-test
## Null hypothesis: no difference
t.test(pre_income ~ program, data = village_randomized)

## graph
plot_diff_preincome <- ggplot(village_randomized, aes(x = program, y = pre_income,
                                                      color = program)) +
  stat_summary(geom = "pointrange", fun.data = "mean_se", fun.args = list(mult = 1.96)) +
  guides(color = FALSE) +
  scale_color_manual(values = c("maroon", "purple"))+
  labs(x = NULL, y = "Pre Income")
plot_diff_preincome
```

There is no significant difference in pre income between those who received the training program (treatment group) and those who did not (control group).  

### 2. Find differences in average outcome in treatment and control groups

You are interested in the causal effect of the program, or
$$E(Postincome|do(Program))$$

You can find this causal effect by calculating the average treatment effect
$$ATE = E(Postincome|Program = 1) - E(Postincome|Program = 0) $$

This is simply the average outcome for people in the pogram minus the average outcome for people not in the program. 

```{r}
village_randomized %>%
  group_by(program) %>%
  summarize(avg_post = mean(post_income))
```
1279.225 - 1179.977 = 99.248 is the causal effect. The program caused an increase in $99 in incomes on average.

You can easily get the same results by running a regression model with post-program income as the outcome variable and the program indicator variable as the explanatory variable. The coefficient for program is the causal effect.

```{r}
model_rct <- lm(post_income ~ program, data = village_randomized)
tidy(model_rct)
```
Based on the RCT, you conclude that the program causes an average increase of $99.25 in income.
We can observe these results using a t-test and/or a graph.

```{r}
## t-test
## Null hypothesis: no difference
t.test(post_income ~ program, data = village_randomized)## we reject the null hypothesis and conclude that there  is a significant difference in post income between the treatment group and control group.

## graph

ggplot(village_randomized, aes(x = program, y = post_income, color = program)) +
  stat_summary(geom = "pointrange", fun.data = "mean_se", fun.args = list(mult = 1.96)) +
  guides(color = FALSE) +
  scale_color_manual(values = c("maroon", "purple"))+
  labs(x = NULL, y = "Post income")
```

